lme-r

### 2025-12-01 notebooks/analysis/05b-mixed-models-R-r1.qmd
revision

```{r}
# data wrangling
library(tidyverse)
# data utilities
library(timetk)
# estimating mixed effects models
library(lme4)
library(merTools)
library(glmmTMB)
library(nlme)
# correlograms
library(ggcorrplot)
library(corrplot)
# data visualisation
library(viridis)
library(ggthemes)
library(ggpubr)
library(zoo)
library(showtext)
library(patchwork)
library(ggrepel)
# display regression equation
library(equatiomatic)
# standardise input variables
  #library(arm)
  # reporting regression results
library(broom.mixed)
library(gtsummary)
library(sjPlot)
library(modelsummary)
```

@@

```{r}
# Set working directory
wd <- '/your/path/to/working/directory/'

@@

```{r}
category <- "rdi"
```

@@

df_stringency <- read.csv(paste0(wd, '/data/inputs/covid-stringency/owid-covid-data.csv')) %>%
  filter(location == "Chile") %>% 
  filter(date %in% dates$date) %>% 
  slice(seq(1, n(), by = 7)) %>% 
  mutate(time = seq(0, n() - 1)) %>%
  select(time, stringency_index)

df <- df %>% left_join(df_stringency, by = "time") %>% rename("O" = "cat")
```

```{r}
df_stringency$stringency_index
```

@@

# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5 <- y ~ time + (time|Of) + (time|Df)
# m5 <- glmmTMB(eq5, data = df)
# m5_re <- plot_model(m5, type ="re")
# 
# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5s <- y ~ time + (time|Of) + (time|Df) + stringency_index
# m5s <- glmmTMB(eq5s, data = df)
# m5s_re <- plot_model(m5s, type ="re")

@@

# eq6 <- y ~ time + time2 + (time|O) + (time|D)
# m6 <- glmmTMB(eq6, data = df)
# m6_re <- plot_model(m6, type ="re")
# 
# eq6s <- y ~ time + time2 + (time|O) + (time|D) + stringency_index
# m6s <- glmmTMB(eq6s, data = df)
# m6s_re <- plot_model(m6s, type ="re")

@@

# eq7 <- y ~ time + time2 + (time + time2|O) + (time + time2|D)
# m7 <- glmmTMB(eq7, data = df)
# m7_re <- plot_model(m7, type ="re")
# 
# eq7s <- y ~ time + time2 + (time + time2|O) + (time + time2|D) + stringency_index
# m7s <- glmmTMB(eq7s, data = df)
# m7s_re <- plot_model(m7s, type ="re")

@@

@@

@@

```{r}
models_o <- list(
  m1,
  m2o,
  m3o,
  m4o,
  m5o,
  m6o,
  m7o
  )
tidy_results_o <- map_df(models_o, tidy, .id = "model")
tidy_results_o <- map_df(models_o, tidy, .id = "model")
file_tidy_results_o <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_o_new.csv')
write.csv(tidy_results_o, file_tidy_results_o, row.names = FALSE)

@@

models_os <- list(
  m1s,
  m2os,
  m3os,
  m4os,
  m5os,
  m6os,
  m7os
  )
tidy_results_os <- map_df(models_os, tidy, .id = "model")
file_tidy_results_os <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_os_new.csv')
write.csv(tidy_results_os, file_tidy_results_os, row.names = FALSE)

@@

file_m7os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m7os_re_new.csv')
write.csv(m7os_re[[1]], file_m7os_re, row.names = FALSE)

@@

```{r}
m3o_re
m3os_re
m4o_re
m4os_re
m5o_re
m5os_re
m6o_re
m6os_re
m7o_re
m7os_re
```

>>>>

```{r}
# data wrangling
library(tidyverse)
# data utilities
library(timetk)
# estimating mixed effects models
library(lme4)
library(merTools)
library(glmmTMB)
library(nlme)
# correlograms
library(ggcorrplot)
library(corrplot)
# data visualisation
library(viridis)
library(ggthemes)
library(ggpubr)
library(zoo)
library(showtext)
library(patchwork)
library(ggrepel)
# display regression equation
library(equatiomatic)
# standardise input variables
  #library(arm)
  # reporting regression results
library(broom.mixed)
library(gtsummary)
library(sjPlot)
library(modelsummary)
library(brms)
```

@@

```{r}
# Set working directory
wd <- '/your/path/to/working/directory/'
wd <- '/Users/carmen/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/research/latin-mobility-covid'

@@

```{r}
category <- "density"
```

@@

df_stringency <- read.csv(paste0(wd, '/data/inputs/covid-stringency/owid-covid-data.csv')) %>%
  filter(location == country) %>% 
  filter(date %in% dates$date) %>% 
  slice(seq(1, n(), by = 7)) %>% 
  mutate(time = seq(0, n() - 1)) %>%
  select(time, stringency_index)

df <- df %>% left_join(df_stringency, by = "time") %>% rename("O" = "cat")

@@

# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5 <- y ~ time + (time|Of) + (time|Df) + ratio
# m5 <- glmmTMB(eq5, data = df)
# m5_re <- plot_model(m5, type ="re")
# 
# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5s <- y ~ time + (time|Of) + (time|Df) + stringency_index + ratio
# m5s <- glmmTMB(eq5s, data = df)
# m5s_re <- plot_model(m5s, type ="re")

@@

# eq6 <- y ~ time + time2 + (time|O) + (time|D) + ratio
# m6 <- glmmTMB(eq6, data = df)
# m6_re <- plot_model(m6, type ="re")
# 
# eq6s <- y ~ time + time2 + (time|O) + (time|D) + stringency_index + ratio
# m6s <- glmmTMB(eq6s, data = df)
# m6s_re <- plot_model(m6s, type ="re")

@@

# eq7 <- y ~ time + time2 + (time + time2|O) + (time + time2|D) + ratio
# m7 <- glmmTMB(eq7, data = df)
# m7_re <- plot_model(m7, type ="re")
# 
# eq7s <- y ~ time + time2 + (time + time2|O) + (time + time2|D) + stringency_index + ratio
# m7s <- glmmTMB(eq7s, data = df)
# m7s_re <- plot_model(m7s, type ="re")

@@

df$O  <- factor(df$O)
eq8o <- y ~ 
    s(time, k = 8) + 
    s(time, O, bs = "fs", k = 7)
m_gam <- gam(eq8o, data = df, method = "REML")

df$O  <- factor(df$O)
eq8os <- y ~ 
    s(time, k = 8) + 
    stringency_index + 
    s(time, O, bs = "fs", k = 7)
m_gams <- gam(eq8os, data   = df, method = "REML")

@@

```{r}
summary(m_gam)$s.table
```

```{r}
plot(m_gam, pages = 1, rug = FALSE, scheme = 1, shade = TRUE)
```

@@

```{r}
models_o <- list(
  m1,
  m2o,
  m3o,
  m4o,
  m5o,
  m6o,
  m7o, 
  m_gam
  )
tidy_fixed <- map_df(models_o, ~ tidy(.x, parametric = TRUE), .id = "model")
tidy_smooth <- map_df(models_o, ~ tidy(.x, smooth = TRUE), .id = "model")

tidy_results_o <- bind_rows(tidy_fixed, tidy_smooth) %>%
  distinct()

file_tidy_results_o <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_o_new.csv')
write.csv(tidy_results_o, file_tidy_results_o, row.names = FALSE)

@@

models_os <- list(
  m1s,
  m2os,
  m3os,
  m4os,
  m5os,
  m6os,
  m7os,
  m_gams
  )
tidy_fixed <- map_df(models_os, ~ tidy(.x, parametric = TRUE), .id = "model")
tidy_smooth <- map_df(models_os, ~ tidy(.x, smooth = TRUE), .id = "model")
tidy_results_os <- bind_rows(tidy_fixed, tidy_smooth) %>%
  distinct()
file_tidy_results_os <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_os_new.csv')
write.csv(tidy_results_os, file_tidy_results_os, row.names = FALSE)

@@

file_m7os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m7os_re_new.csv')
write.csv(m7os_re[[1]], file_m7os_re, row.names = FALSE)
```

```{r}
# m3o_re
# m3os_re
# m4o_re
# m4os_re
# m5o_re
# m5os_re
# m6o_re
# m6os_re
# m7o_re
# m7os_re
```

## Compute recovery times for spline model. The 95% confidence intervals are computed via bootstrap.

```{r}
stringency <- FALSE
```

```{r}
# Create new time sequence covering the range of your original data
new_time <- expand.grid(
  time = seq(min(df$time), max(df$time)*100, length.out = max(df$time)*100-min(df$time)+1),
  O = levels(df$O)   # one dataset per group
)

new_time <- new_time %>%
  left_join(df[, c("time", "O", "stringency_index")], by = c("time", "O"))

new_time$stringency_index[is.na(new_time$stringency_index)] <- 0

if (stringency == FALSE) {
  new_time <- expand.grid(
  time = seq(min(df$time), max(df$time)*100, length.out = max(df$time)*100-min(df$time)+1),
  O = levels(df$O)   # one dataset per group
)
  pred_base <- predict(m_gam, newdata = new_time, se.fit = TRUE)
} else {
  pred_base <- predict(m_gams, newdata = new_time, se.fit = TRUE)
}

new_time$fit <- pred_base$fit
new_time$se  <- pred_base$se.fit

# Find FIRST root where preds(t) == target
get_root_threshold <- function(time, preds, target) {
  diff_sign <- diff(sign(preds - target))
  crossings <- which(diff_sign != 0)
  if (length(crossings) == 0) return(NA)

  i <- crossings[1]  # first crossing
  out <- tryCatch(
    uniroot(function(t) {
      approx(time, preds - target, xout = t)$y
    },
    lower = time[i],
    upper = time[i + 1])$root,
    error = function(e) NA
  )
  return(out)
}

alphas <- seq(0, 1, by = 0.05)   # adjust resolution as needed

# Bootstrap confidence intervals
set.seed(123)
n_boot <- 200   # increase to 500–1000 for publication

results_alpha <- map_df(levels(df$O), function(g) {

  group_data <- new_time %>% filter(O == g)

  # group-specific y(t=0): use earliest time in grid for that group
  idx0 <- which.min(group_data$time)
  y0_group <- group_data$fit[idx0]

  # For each alpha / target, bootstrap
  map_df(seq_along(alphas), function(j) {
    alpha   <- alphas[j]
    target <- (1 - alpha) * y0_group

    boot_roots <- replicate(n_boot, {
      # simulate one curve from uncertainty
      y_sim <- rnorm(nrow(group_data), mean = group_data$fit, sd = group_data$se)
      get_root_threshold(group_data$time, y_sim, target)
    })

    data.frame(
      O        = g,
      alpha    = alpha,
      y_target = target,
      t_alpha  = median(boot_roots, na.rm = TRUE),
      CI_lower = quantile(boot_roots, 0.025, na.rm = TRUE),
      CI_upper = quantile(boot_roots, 0.975, na.rm = TRUE),
      SD_t_alpha = sd(boot_roots, na.rm = TRUE)
    )
  })
})

results_alpha <- as_tibble(results_alpha)

@@

## Save recovery times for splines model

if (stringency == FALSE) {
  file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-gam-r1.csv')
} else {
  file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-gams-r1.csv')
}
write.csv(results_alpha, file_results_alpha, row.names = FALSE)

## Plot recovery times splines model for various alpha

```{r}
ggplot(results_alpha, aes(x = alpha, y = t_alpha, color = O, group = O)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper, fill = O),
              alpha = 0.15, color = NA) +
  labs(
    title = "Time to reach threshold level (1 - α)",
    x = expression(alpha),
    y = expression(t(alpha)),
    color = "Group (O)",
    fill = "Group (O)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    plot.title = element_text(face = "bold", size = 16)
  )

### 2025-12-01 notebooks/analysis/08b-mixed-models-R-cities-r1.qmd
revision

```{r}
category <- "density"
capital <- TRUE

@@

@@

```{r}
models_cat <- list(
  m1,
  m2cat,
  m3cat,
  m4cat,
  m5cat,
  m6cat,
  m7cat
  )
tidy_results_cat <- map_df(models_cat, tidy, .id = "model")
file_tidy_results_cat <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_tidy_results_cat.csv')
write.csv(tidy_results_cat, file_tidy_results_cat, row.names = FALSE)

glance_results_cat <- map_df(models_cat, glance, .id = "model")
file_glance_results_cat <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_glance_results_cat.csv')
write.csv(glance_results_cat, file_glance_results_cat, row.names = FALSE)

models_cats <- list(
  m1s,
  m2cats,
  m3cats,
  m4cats,
  m5cats,
  m6cats,
  m7cats
  )
tidy_results_cats <- map_df(models_cats, tidy, .id = "model")
file_tidy_results_cats <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_tidy_results_cats.csv')
write.csv(tidy_results_cats, file_tidy_results_cats, row.names = FALSE)

glance_results_cats <- map_df(models_cats, glance, .id = "model")
file_glance_results_cats <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_glance_results_cats.csv')
write.csv(glance_results_cats, file_glance_results_cats, row.names = FALSE)
```

@@

file_m3cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m3cat_re.csv')
write.csv(m3cat_re[[1]], file_m3cat_re, row.names = FALSE)

file_m3cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m3cats_re.csv')
write.csv(m3cats_re[[1]], file_m3cats_re, row.names = FALSE)

@@

file_m4cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m4cat_re.csv')
write.csv(m4cat_re[[1]], file_m4cat_re, row.names = FALSE)

file_m4cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m4cats_re.csv')
write.csv(m4cats_re[[1]], file_m4cats_re, row.names = FALSE)

@@

file_m5cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m5cat_re.csv')
write.csv(m5cat_re[[1]], file_m5cat_re, row.names = FALSE)

file_m5cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m5cats_re.csv')
write.csv(m5cats_re[[1]], file_m5cats_re, row.names = FALSE)

@@

file_m6cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m6cat_re.csv')
write.csv(m6cat_re[[1]], file_m6cat_re, row.names = FALSE)

file_m6cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m6cats_re.csv')
write.csv(m6cats_re[[1]], file_m6cats_re, row.names = FALSE)

@@

file_m7cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m7cat_re.csv')
write.csv(m7cat_re[[1]], file_m7cat_re, row.names = FALSE)

file_m7cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m7cats_re.csv')
write.csv(m7cats_re[[1]], file_m7cats_re, row.names = FALSE)

@@

# Plot using ggplot
ggplot() +
  geom_point(data = df, aes(x = time, y = y), color = "blue", alpha = 0.6, size = 2) +
  geom_point(data = pred_df_1, aes(x = time, y = fitted), color = "red", alpha = 0.6, size = 1.5) +
  geom_point(data = pred_df_2, aes(x = time, y = fitted), color = "red3", alpha = 0.6, size = 1.5) +
  geom_point(data = pred_df_3, aes(x = time, y = fitted), color = "darkred", alpha = 0.6, size = 1.5) +
  geom_point(data = pred_df_4, aes(x = time, y = fitted), color = "purple4", alpha = 0.6, size = 1.5) +
  geom_point(data = pred_df_5, aes(x = time, y = fitted), color = "midnightblue", alpha = 0.6, size = 1.5) +
  labs(title = "Observed vs Predicted (Quadratic Fit)",
       x = "Time", y = "Y") +
  xlim(0, 300) +
  ylim(min(c(df$y, pred_df$fitted), na.rm = TRUE),
       max(c(df$y, pred_df$fitted, 20), na.rm = TRUE)) +
  theme_minimal()

@@

>>>>

```{r}
category <- "density"
capital <- FALSE

@@

df$cat  <- factor(df$cat)
eq8cat <- y ~ 
    s(time, k = 8) + 
    s(time, cat, bs = "fs", k = 7)
m_gam <- gam(eq8cat, data = df, method = "REML")

df$cat  <- factor(df$cat)
eq8cats <- y ~ 
    s(time, k = 8) + 
    stringency_index + 
    s(time, cat, bs = "fs", k = 7)
m_gams <- gam(eq8cats, data   = df, method = "REML")

@@

```{r}
models_cat <- list(
  m1,
  m2cat,
  m3cat,
  m4cat,
  m5cat,
  m6cat,
  m7cat, 
  m_gam
  )
tidy_fixed <- map_df(models_cat, ~ tidy(.x, parametric = TRUE), .id = "model")
tidy_smooth <- map_df(models_cat, ~ tidy(.x, smooth = TRUE), .id = "model")
tidy_results_cat <- bind_rows(tidy_fixed, tidy_smooth) %>%
  distinct()
file_tidy_results_cat <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_tidy_results_cat_new.csv')
# write.csv(tidy_results_cat, file_tidy_results_cat, row.names = FALSE)

glance_results_cat <- map_df(models_cat, glance, .id = "model")
file_glance_results_cat <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_glance_results_cat_new.csv')
# write.csv(glance_results_cat, file_glance_results_cat, row.names = FALSE)

models_cats <- list(
  m1s,
  m2cats,
  m3cats,
  m4cats,
  m5cats,
  m6cats,
  m7cats,
  m_gams
  )
tidy_fixed <- map_df(models_cats, ~ tidy(.x, parametric = TRUE), .id = "model")
tidy_smooth <- map_df(models_cats, ~ tidy(.x, smooth = TRUE), .id = "model")
tidy_results_cats <- bind_rows(tidy_fixed, tidy_smooth) %>%
  distinct()

file_tidy_results_cats <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_tidy_results_cats_new.csv')
# write.csv(tidy_results_cats, file_tidy_results_cats, row.names = FALSE)

glance_results_cats <- map_df(models_cats, glance, .id = "model")
file_glance_results_cats <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_netflows_', place, '_param_glance_results_cats_new.csv')
# write.csv(glance_results_cats, file_glance_results_cats, row.names = FALSE)
```

@@

file_m3cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m3cat_re_new.csv')
# write.csv(m3cat_re[[1]], file_m3cat_re, row.names = FALSE)

file_m3cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m3cats_re_new.csv')
# write.csv(m3cats_re[[1]], file_m3cats_re, row.names = FALSE)

@@

file_m4cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m4cat_re_new.csv')
# write.csv(m4cat_re[[1]], file_m4cat_re, row.names = FALSE)

file_m4cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m4cats_re_new.csv')
# write.csv(m4cats_re[[1]], file_m4cats_re, row.names = FALSE)

@@

file_m5cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m5cat_re_new.csv')
# write.csv(m5cat_re[[1]], file_m5cat_re, row.names = FALSE)

file_m5cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m5cats_re_new.csv')
# write.csv(m5cats_re[[1]], file_m5cats_re, row.names = FALSE)

@@

file_m6cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m6cat_re_new.csv')
# write.csv(m6cat_re[[1]], file_m6cat_re, row.names = FALSE)

file_m6cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m6cats_re_new.csv')
# write.csv(m6cats_re[[1]], file_m6cats_re, row.names = FALSE)

@@

file_m7cat_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m7cat_re_new.csv')
# write.csv(m7cat_re[[1]], file_m7cat_re, row.names = FALSE)

file_m7cats_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_', place, '_m7cats_re_new.csv')
# write.csv(m7cats_re[[1]], file_m7cats_re, row.names = FALSE)

@@

@@

## Compute recovery times for spline model. The 95% confidence intervals are computed via bootstrap.

```{r}
stringency <- FALSE
```

```{r}
# Create new time sequence covering the range of your original data
new_time <- expand.grid(
  time = seq(min(df$time), max(df$time)*100, length.out = max(df$time)*100-min(df$time)+1),
  cat = levels(df$cat)   # one dataset per group
)

new_time <- new_time %>%
  left_join(df[, c("time", "cat", "stringency_index")], by = c("time", "cat"))

new_time$stringency_index[is.na(new_time$stringency_index)] <- 0

if (stringency == FALSE) {
  new_time <- expand.grid(
  time = seq(min(df$time), max(df$time)*100, length.out = max(df$time)*100-min(df$time)+1),
  cat = levels(df$cat)   # one dataset per group
)
  pred_base <- predict(m_gam, newdata = new_time, se.fit = TRUE)
} else {
  pred_base <- predict(m_gams, newdata = new_time, se.fit = TRUE)
}

new_time$fit <- pred_base$fit
new_time$se  <- pred_base$se.fit

# Find FIRST root where preds(t) == target
get_root_threshold <- function(time, preds, target) {
  diff_sign <- diff(sign(preds - target))
  crossings <- which(diff_sign != 0)
  if (length(crossings) == 0) return(NA)

  i <- crossings[1]  # first crossing
  out <- tryCatch(
    uniroot(function(t) {
      approx(time, preds - target, xout = t)$y
    },
    lower = time[i],
    upper = time[i + 1])$root,
    error = function(e) NA
  )
  return(out)
}

alphas <- seq(0, 1, by = 0.05)   # adjust resolution as needed

# Bootstrap confidence intervals
set.seed(123)
n_boot <- 200   # increase to 500–1000 for publication

results_alpha <- map_df(levels(df$cat), function(g) {

  group_data <- new_time %>% filter(cat == g)

  # group-specific y(t=0): use earliest time in grid for that group
  idx0 <- which.min(group_data$time)
  y0_group <- group_data$fit[idx0]

  # For each alpha / target, bootstrap
  map_df(seq_along(alphas), function(j) {
    alpha   <- alphas[j]
    target <- (1 - alpha) * y0_group

    boot_roots <- replicate(n_boot, {
      # simulate one curve from uncertainty
      y_sim <- rnorm(nrow(group_data), mean = group_data$fit, sd = group_data$se)
      get_root_threshold(group_data$time, y_sim, target)
    })

    data.frame(
      cat        = g,
      alpha    = alpha,
      y_target = target,
      t_alpha  = median(boot_roots, na.rm = TRUE),
      CI_lower = quantile(boot_roots, 0.025, na.rm = TRUE),
      CI_upper = quantile(boot_roots, 0.975, na.rm = TRUE),
      SD_t_alpha = sd(boot_roots, na.rm = TRUE)
    )
  })
})

results_alpha <- as_tibble(results_alpha)

## Save recovery times for splines model

if (stringency == FALSE) {
  if (capital == TRUE) {
    file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-capital-gam-r1.csv')
  } else {
    file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-fuas-gam-r1.csv')
  }
} else {
  if (capital == TRUE) {
    file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-capital-gams-r1.csv')
  } else {
    file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-fuas-gams-r1.csv')
  }
}
write.csv(results_alpha, file_results_alpha, row.names = FALSE)

## Plot recovery times splines model for various alpha

```{r}
ggplot(results_alpha, aes(x = alpha, y = t_alpha, color = cat, group = cat)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper, fill = cat),
              alpha = 0.15, color = NA) +
  labs(
    title = "Time to reach threshold level (1 - α)",
    x = expression(alpha),
    y = expression(t(alpha)),
    color = "Group (cat)",
    fill = "Group (cat)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    plot.title = element_text(face = "bold", size = 16)
  )
